---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Coffee

The **coffee** package provides tools and interactive features to analyze coffee consumption patterns based on survey data. This package includes a variety of functions for statistical analysis and an RShiny app to help users visualize coffee preferences by demographic factors, such as age group, gender, and favorite coffee type.

Explore coffee preferences interactively using the Shiny app included with this package, allowing you to filter data and view results in various visualizations tailored to your preferences.

You can view a comprehensive guide to this package on its website.

**NOTE**: When launching the RShiny app, there may be a short delay as the app loads the data. Please allow the app approximately 10-20 seconds to fully initialize, as it loads and processes the coffee survey dataset.

## Installation

You can install the development version of the **Coffee** package from GitHub with:

```{r github,eval=FALSE}
devtools::install_github("azham3033/Coffee")
```



This example provides a bar plot showing the average daily coffee consumption by age group based on the `coffee` dataset.

## Shiny App

The `coffee` package also includes an interactive Shiny app for exploring coffee survey data. To launch the app, use the following command:

```{r shinyapp, eval=FALSE}
library(Coffee)
launch_app()
```

The app enables you to:

Filter the data by age group, gender, and favorite coffee type. View visualizations of coffee preferences segmented by demographics. Discover trends and insights into coffee consumption patterns.

------------------------------------------------------------------------
